---
title: "Agriculture Emissions Data Wrangling"
author: "Ben DeMoras"
date: "2/8/2022"
output: 
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Needed:

+ N2O emissions (Response variable)

  + Broken down into synthetic fertilizers, manure applied on soils, and N2O emissions. _Manure applied on soils_ is what farmers are actively applying as fertilizer, whereas _left on soils_ is the manure deposited by grazing livestock. I only include _applied to soils_ because you don't have synthetic-pooping livestock.
  
  + Filter down to N2O emissions from synthetic and total manure.
  
  + Source: See below.

+ Manure and Synthetic Fertilizer use: (Treatment)

  + Filter down to tons of synthetic fertilizer and tons manure applied 
  
    + Synthetic: https://www.fao.org/faostat/en/#data/GY
    
    + Manure (applied to soil):  https://www.fao.org/faostat/en/#data/GU
    
  + Filter agricultural cropland area data https://www.fao.org/faostat/en/#data/RL
  
  + Calculate manure/fertilizer use per acre
  
## Import Data
I am Using Tidyverse packages to more easily import data to R and filter down what we need. Cheatsheet: https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf
```{r message=FALSE, warning=FALSE}
# install.packages('dplyr')
  library(dplyr)

# 1. Import Data
manure.csv <- read.csv(file="csv-data/Emissions_Manure.csv", header = TRUE)
synth.csv  <- read.csv(file="csv-data/Emissions_Synthetic.csv", header = TRUE)
```


## Tidy Data
NOTE: FAO breaks down fertilizers according to how the emissions end up. We want to consider these as one source.
**We need to be sure we are combining the right categories!**
```{r echo = TRUE, results = 'hide'}
# 2. Define some variables we'll need later on.
manure.use.filter = c(
    "Manure applied to soils (N content)",
    "Manure applied to soils that leaches (N content)",
    "Manure applied to soils that volatilises (N content)")

manure.emit.filter = c(
    "Direct emissions (N2O) (Manure applied)",
    "Indirect emissions (N2O that leaches) (Manure applied)",
    "Indirect emissions (N2O that volatilises) (Manure applied)",
    "Indirect emissions (N2O) (Manure applied)")

synth.use.filter = c(
    "Agricultural Use in nutrients",
    "Nitrogen fertilizer content applied that leaches",
    "Nitrogen fertilizer content applied that volatilises")

synth.emit.filter = c(
    "Direct emissions (N2O) (Synthetic fertilizers)",
    "Indirect emissions (N2O that leaches) (Synthetic fertilizers)",
    "Indirect emissions (N2O that volatilises) (Synthetic fertilizers)",
    "Indirect emissions (N2O) (Synthetic fertilizers)")

# 3. Get only the data we need
manure.use <- manure.csv %>% 
    filter(Element %in% manure.use.filter) %>%
    select(Area, Y2019) %>% 
    # Manure data are broken down by animal so now we need to combine them by country
    group_by(Area) %>% 
    summarize(Kg.ManureN.2019 = sum(Y2019, na.rm = TRUE))

manure.emit <- manure.csv %>% 
    filter(Element %in% manure.emit.filter) %>% 
    select(Area, Y2019) %>%
    group_by(Area) %>% 
    summarize(KT.ManureN2O.2019 = sum(Y2019, na.rm = TRUE))

synth.use <- synth.csv %>%
    filter(Element %in% synth.use.filter) %>% 
    select(Area, Y2019) %>% 
    group_by(Area) %>% 
    summarize(Kg.SynthNutrient.2019 = sum(Y2019, na.rm = TRUE))

synth.emit <- synth.csv %>% 
    filter(Element %in% synth.emit.filter) %>% 
    select(Area, Y2019) %>% 
    group_by(Area) %>% 
    summarize(KT.SynthN20.2019 = sum(Y2019, na.rm = TRUE))

# 4. Display subset of results in console to confirm success
# head(manure.use)
# head(manure.emit)
# head(synth.use)
# head(synth.emit)

```


## Combine the data
Now, we're ready to combine the data into one dataframe for analysis!
```{r}
data.master <- manure.use %>% 
    inner_join(manure.emit, by = "Area") %>% 
    inner_join(synth.use,   by = "Area") %>% 
    inner_join(synth.emit,  by = "Area")

head(data.master, n = 10L)
```

We still need to exclude some categories (ie, `Annex I countries`, `Africa`) and figure out if we'll analyze the whole dataset, or a random subset.